Himalayan Expeditions Success Analysis 

### 2024-04-14 paper/paper.qmd
Introduction added

# Introduction

1) broader context to motivate; 
2) some detail about what the paper is about; 
3) a clear gap that needs to be filled; 
4) what was done; 
5) what was found; 
6) why it is important;
7) Estimand

(Likely 3 or 4 paragraphs, or 10 per cent of total.)

@@

# Model {#sec-model}
I used a Bayesian Logistic Regression model to find the probability that someone will successfully complete the Himalayan peak they are on the expedition for. Logistic regression is a method used for binary classification to predict the probability of a categorical dependent variable.

For my analysis, a logistic regression model will be used to model if the person will be successful on their attempt to climb the particular Himalayan peak. The model will be based on five independent demographic variables: `height of the peak`, `sex`, `age`, `seasons` and `solo`.

@@

- $\hat{p}$ represents the probability that someone will successfully complete the peak they are on the expedition for.
- $\beta_0$ represents the intercept term of this logistical regression. It is the probability that someone will successfully complete the peak they are on the expedition for if the predictors' values are zero
- $\beta_1$ is the coefficient corresponding to height of the peak
- $\beta_2$ is the coefficient corresponding to sex
- $\beta_3$ is the coefficients corresponding to age
- $\beta_4$ is the coefficients corresponding to seasons
- $\beta_5$ is the coefficients corresponding to solo

@@

# Create a data frame with all combinations of sex and seasons
sex_levels <- c('M', 'F')  # Replace with actual categories
seasons_levels <- unique(analysis_data$seasons)  # Extract unique seasons from your analysis data

new_data <- expand.grid(
  sex = sex_levels,
  seasons = seasons_levels
)

# Predicting success based on sex and seasons
predictions <- predict(expeditions_model, newdata = new_data, type = "response")

# Create a data frame to display the predictions
prediction_df <- data.frame(
  sex = new_data$sex, 
  seasons = new_data$seasons, 
  success_probability = predictions
)

>>>>

# Introduction
Mountaineering, with its blend of adventure and challenge, has captivated explorers for generations. It represents a pinnacle of human endeavor, testing physical prowess, mental fortitude, and strategic planning against some of the most formidable natural landscapes on Earth. Amidst the allure of conquering these majestic peaks, lies a critical question: what factors contribute to the likelihood of a successful summit attempt? This paper delves into precisely this inquiry.

There are numerous mountain ranges around the world where people embark on expeditions. One of the most famous mountain ranges are the Himalayas, which also consist of the highest mountain peak in the world, Mt. Everest. Our paper uses the data from expeditions to the Himalayan mountain ranges in Nepal.

Our estimand is the relationship between different demographic, environmental and geographic factors (such as height of the peak, sex, age, season of expedition and if it was a solo ascent) and the successfulness of climbing a summit. Using our extensive analysis dataset, our goal is to identify trends and factors that influence a successful expedition and eventually conclude what factors help in a more successful attempt.

We use data from Alex Cookson’s datasets (link) which are sourced from The Himalayan Expedition records (link), to understand these factors and trends. This is done by leveraging a Bayesian Logistic Regression model and then predicting the probability of a successful attempt over various demographic and environmental factors.

Our analysis led to us … (add what was found)

This research highlights the interplay between climber attributes and environmental conditions, offering insights into Himalayan expeditions. These findings have practical implications for expedition planning, risk management, and safety protocols. Ultimately, our study aims to improve decision-making in high-altitude mountaineering, making it safer and more informed in one of the world's most challenging environments.

@@

# Model {#sec-model}
I used a Bayesian Logistic Regression model to find the probability that someone will successfully summit the Himalayan peak they are on the expedition for. Logistic regression is a method used for binary classification to predict the probability of a categorical dependent variable.

For my analysis, a logistic regression model will be used to model if the person will be successful on their attempt to summit the particular Himalayan peak. The model will be based on five independent demographic variables: `height of the peak`, `sex`, `age`, `seasons` and `solo`.

@@

- $\hat{p}$ represents the probability that someone will successfully complete the peak they are on the expedition for.
- $\beta_0$ represents the intercept term of this logistical regression. It is the probability that someone will successfully complete the peak they are on the expedition for if the predictors' values are zero
- $\beta_1$ is the coefficient corresponding to height of the peak
- $\beta_2$ is the coefficient corresponding to sex of the person
- $\beta_3$ is the coefficients corresponding to age of the person
- $\beta_4$ is the coefficients corresponding to seasons of the person
- $\beta_5$ is the coefficients corresponding to if the person is attempting the summit alone.

@@
